A GPU Sonar Simulator for Automatic Target Recognition

### 2014-04-30 UAM2014/Article/article.tex
la2pdf commit - 2014:04:30	23:33:59	(Wed30.Apr)

Various simulators for high frequency, side-looking sonar imagery have been published, e.g.~\cite{Bell1997,Sammelmann2004}. Most implementations have however prioritized accurate acoustic modeling at the expense of execution speed. One exception is the SIGMAS+ simulator in~\cite{Coiras2009a, Coiras2009b} which, similar to our approach, takes advantage of parallel processing at GPUs and uses OpenGL to render a 2D image from a 3D model, under the assumption of Lambertian scattering. However, SIGMAS+ is still aimed more towards realistically looking images, including effects like ambient noise, which is irrelevant for our purpose of template generation. Also, their simulator sums images obtained from multiple OpenGL rendering passes to create the sonar image, while we render once and post-process this result with OpenCL to tailor the sonar image with more flexibility.

@@

\newlength\imgspacing\setlength\imgspacing{.5cm}

Whenever conventional computer graphics processing can be used to 
GPUs are designed to provide the best possible graphics processing performance, and there large ecosystem of tools and libraries  available 
Using a GPU to drive a simulator is advantageous since

Using a GPU-based simulator is attractive since computer is an attractive solution as there are a  graphics traits. For one over a CPU-based standard simulator is two-fold: First is its ability to create templates incredibly fast. This makes it possible to tailor the templates more closely to an image as the parameters needed by the simulator can be directly estimated from the image. In a standard approach, on the other hand, a template library must be created beforehand from a limited set of parameter values. Hence, it may happen that no template in the library is a close fit of the object in the image even though the object actually is a target of interest.

@@

\setcounter{topnumber}{2}

@@

One way to automatically classify objects in SAS images is compare the imaged objects with a predefined set of templates. However, this is suboptimal as it is infeasible to create templates for all configurations of seabeds and objects. Instead we have implemented a SAS simulator that creates the templates in delayed real-time based on parameters estimated from the current scene. In our studies this improves the match between the SAS objects and the template considerably.

To obtain the delayed real-time performance we wanted to use modern GPUs. Their theoretical peak performance is typically an order of magnitude higher than a CPU in a comparable price range. While it is hard to obtain this performance in practice, we got very close by using a well matured and highly optimized graphics processing API such as OpenGL. Because of this we use OpenGL for rendering a 2D projection of a scene with 3D models of the seabed and objects. This is finally converted using the OpenCL language, which can interoperate with OpenGLTo convert this into a sonar template image we use Some additional prost-processing is required as well to form the actual template, but

>>>>

Various simulators for high frequency, side-looking sonar imagery have been published, e.g.~\cite{Bell1997,Sammelm2003}. Most implementations have however prioritized accurate acoustic modeling at the expense of execution speed. One exception is the SIGMAS+ simulator in~\cite{Coiras2009a, Coiras2009b} which, similar to our approach, takes advantage of parallel processing at GPUs and uses OpenGL to render a 2D image from a 3D model, under the assumption of Lambertian scattering. However, SIGMAS+ is still aimed more towards realistically looking images, including effects like ambient noise, which is irrelevant for our purpose of template generation. Also, their simulator sums images obtained from multiple OpenGL rendering passes to create the sonar image, while we render once and post-process this result with OpenCL to tailor the sonar image with more flexibility.

@@

\newlength\imgspacing\setlength\imgspacing{.5cm}
\setcounter{topnumber}{2}

% Whenever conventional computer graphics processing can be used to 
% GPUs are designed to provide the best possible graphics processing performance, and there large ecosystem of tools and libraries  available 
% Using a GPU to drive a simulator is advantageous since
% 
% 
% Creating a GPU-based simulator is attractive since computer is an attractive solution as there are a  graphics traits. For one over a CPU-based standard simulator is two-fold:

Our GPU-based simulator is very fast. On a computer with a six-core Intel Core i7-3930K and a Radeon HD7970 it computes almost 1000 templates per second with a size of 1000x1000 pixels (1 megapixel). This makes it possible to tailor the templates more closely to an image as the parameters needed by the simulator can be directly estimated from the image. In a standard approach, on the other hand, a template library must be created beforehand from a limited set of parameter values. Hence, it may happen that no template in the library is a close fit of the object in the image even though the object actually is a target of interest.

@@

@@

One way to automatically classify objects in SAS images is compare the imaged objects with a predefined set of templates. However, this is suboptimal as it is infeasible to create templates for all configurations of seabeds and objects. Instead we have implemented a SAS simulator that creates the templates in delayed real-time based on parameters estimated from the current scene. In our studies this improves the match between the SAS objects and the template significantly in most cases.

To obtain the delayed real-time performance we implemented the simulator on a GPU. Their theoretical peak performance is typically an order of magnitude higher than a CPU in a comparable price range. This potential is effectively utilized with a well matured and highly optimized graphics processing API such as OpenGL. Our simulator use OpenGL for most for most of the scene processing. A final post-processing step is performed in OpenCL

\begin{itemize}
\item Mer testing
\item Robustifisering
\item Tuning
\end{itemize}
